Coming to Germany: Language, Job search and Labor market outcomes

### 2018-05-08 Paper 2.tex
Update Paper 2.tex

Empirical evidence in the existing literature on language skills of immigrants often reports the 'total effect' of language fluency on a range of labor market outcomes. But it is not empirically clear \textbf{why} such an effect arises. This study aims to investigate the role of the pre-migration language skills on the use of social networks to find jobs, which in turn affects current earnings and the quality of jobs. In doing so, I attempt to open the black-box to disentangle how and why language proficiency exerts its influence on labor market success of immigrants. I focus on pre-migration language skills because immigrants who master the national language at the very beginning of migration could immediately find better jobs and obtain an initial advantage which puts them well ahead of other immigrants who start learning the language after migration. Further, by focusing on pre-migration language skills, I am able to decompose the total effect of language skills into indirect effect through preferred early job search channels and direct effect on subsequent labor market success.

@@

\textbf{Independent Variables}

@@

$$Y_{i} = \alpha_{1} + \beta_{1}L_{i} + \xi_{1}X_{i} + \epsilon_{i1} (1)$$
$$M_{i} = \alpha_{2} + \beta_{2}L_{i} + \xi_{2}X_{i} + \epsilon_{i2} (2) $$
$$Y_{i} = \alpha_{3} + \beta_{3}L_{i} + \gamma M_{i} + \xi_{3}X_{i} + \epsilon_{i3} (3)$$

>>>>

Empirical evidence in the existing literature on language skills of immigrants often reports the 'total effect' of language fluency on a range of labor market outcomes. But it is, in some cases, not empirically clear \textbf{why} such an effect arises. This study aims to investigate the role of the pre-migration language skills on the use of social networks to find jobs, which in turn affects current earnings and the quality of jobs. In doing so, I attempt to open the black-box to disentangle how and why language proficiency exerts its influence on labor market success of immigrants. I focus on pre-migration language skills because immigrants who master the national language at the very beginning of migration could immediately find better jobs and obtain an initial advantage which puts them well ahead of other immigrants who start learning the language after migration. Further, by focusing on pre-migration language skills, I am able to decompose the total effect of language skills into indirect effect through preferred early job search channels and direct effect on subsequent labor market success.

@@

\textbf{Independent Variables}
To measure the pre-migration language fluency of immigrants in Germany, I use the self-assessed language skills and the interaction of the linguistic distance between German and a respective mother tongue of an immigrants with his/her age at arrival to Germany. First, the IAB-SOEP Migration Sample includes three questions to ask about the language skills of immigrants at the very beginning of their migration (the so-called pre-migration language skills) with respect to speaking, writing and reading skills. Immigrants self-assessed these skills on the scale of 1-5 (1: Not all all, 2: Poor, 3: Fair, 4: Good, 5: Very good). I use Principal Component Analysis (PCA) where a single latent component (overall German language skills at migration) is captured by seeking a linear combinations of variables (Speaking, Writing, Reading skills) such that the maximum variance is extracted from these variables. Table 1 presents the factor loading or the correlation coefficients between three original variables and the principal component. All three skills are highly correlated with one another and contribute roughly the same to a single latent factor. The resulting prediction of this PCA is a new composite German language proficiency which captures characteristics of all three individual skills, and is standardized to reflect z-score. In other words, the new German language proficiency variable has mean 0 and standard deviation 1. This is our variable of interest, and I will use this variable as a benchmark for comparison with the proxy variable presented below.

Second, I create a proxy variable

@@

\begin{cases} Y_{i} = \alpha_{1} + \beta_{1}L_{i} + \xi_{1}X_{i} + \epsilon_{i1} (1) \\ M_{i} = \alpha_{2} + \beta_{2}L_{i} + \xi_{2}X_{i} + \epsilon_{i2} (2) \\ Y_{i} = \alpha_{3} + \beta_{3}L_{i} + \gamma M_{i} + \xi_{3}X_{i} + \epsilon_{i3} (3)
\end{cases}
